---
title: "PART 2"
author: "Rachel Juma"
date: "2/3/2022"
output: html_document
---
## 1.) Problem Definition

The brand’s Sales and Marketing team of Kira Plastinina,a Russian fashion brand would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups by performing clustering and stating insights drawn from the analysis and visualizations.


## 2.)Data Sourcing

- Dataset is exctrated from "http://bit.ly/EcommerceCustomersDataset".
- It consists of 10 numerical and 8 categorical attributes. The 'Revenue' attribute will be used as the class label.
- "Administrative", "Administrative Duration", "Informational", "Informational Duration", "Product Related" and "Product Related Duration" represents the number of different types of pages visited by the visitor in that session and total time spent in each of these page categories. The values of these features are derived from the URL information of the pages visited by the user and updated in real-time when a user takes an action, e.g. moving from one page to another.

- The "Bounce Rate", "Exit Rate" and "Page Value" features represent the metrics measured by "Google Analytics" for each page in the e-commerce site. 

-The value of the "Bounce Rate" feature for a web page refers to the percentage of visitors who enter the site from that page and then leave ("bounce") without triggering any other requests to the analytics server during that session. 

- The value of the "Exit Rate" feature for a specific web page is calculated as for all pageviews to the page, the percentage that was the last in the session.
- The "Page Value" feature represents the average value for a web page that a user visited before completing an e-commerce transaction. 
- The "Special Day" feature indicates the closeness of the site visiting time to a specific special day (e.g. Mother’s Day, Valentine's Day) in which the sessions are more likely to be finalized with the transaction. The value of this attribute is determined by considering the dynamics of e-commerce such as the duration between the order date and delivery date. For example, for Valentina’s day, this value takes a nonzero value between February 2 and February 12, zero before and after this date unless it is close to another special day, and its maximum value of 1 on February 8. 
- The dataset also includes the operating system, browser, region, traffic type, visitor type as returning or new visitor, a Boolean value indicating whether the date of the visit is weekend, and month of the year.

## 3.)Check the Data

```{r}

library(data.table)
df <- fread("http://bit.ly/EcommerceCustomersDataset")

head(df)     # Preview the data
dim(df)      # shows the size of the data
class(df)    
names(df) # Determine variable names
str(df)
summary(df)
```

- Dataset had 12330 rows and 18 variables. 8 out of 18 are categorical while 10 are numerical.
- The 'Revenue' attribute will be used as the class label. There are 10422 cases of False revenue compared to 1908 cases of True revenue.

## 4.)Perform Data Cleaning

```{r}
# install.packages("lattice")
library(lattice)
# check for null values
colSums(is.na(df))  
df1 <- na.omit(df) # new dataframe with removed null values
str(df1)
dim(df1)
colSums(is.na(df1)) 

# converting logical into factor

df1$Revenue <- as.factor(df1$Revenue) # converting logical into factor 1(FALSE) and 2(TRUE)
df1$Weekend <- as.factor(df1$Weekend)# converting logical into factor 1(FALSE) and 2(TRUE)
df1$VisitorType <- as.factor((df1$VisitorType))
str(df1)
head(df1)

# Using boxplot to check for outliers of numerical variables

cols <- names(df1)[sapply(df1, is.numeric)]
list_diag <- vector("list", length(cols))
                    
for(i in list_diag)
list_diag <- lapply(cols, function(x) 
                    bwplot(reformulate(x), data = df1, main = x))
list_diag

# Using unique function we check for anomalies in the categorical variables.
uniq.Revenue <- unique(df1$Revenue)
uniq.Revenue

uniq.Weekend<-unique(df1$Weekend)
uniq.Weekend

# determining duplicated values
duplicated <- df1[duplicated(df1),]
duplicated

# removing duplicates
df2 <- unique(df1) # new cleaned dataframe
df2
str(df2)
```

- Administrative ,Administrative_Duration, Informational, Informational_Duration, ProductRelated ,ProductRelated_Duration, BounceRates, ExitRates variables have 14 missing values. After dropping them the dataset remains with 12316 rows.
- Dataset has 117 duplicates,so we'll drop them bringing the number of remaining rows to 
12199.

## 5.)Perform Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)
### 5.1.) Univariate
```{r}
# Plotting bar graphs

# 1. Revenue variable
rev <- df2$Revenue
rev.freq <- table(rev)
barplot(rev.freq,col = "magenta",xlab = "Revenue",ylab = "Count")

# 2. Weekend variable
week <- df2$Weekend
week.freq <- table(week)
barplot(week.freq,col = "green",xlab = "Weekend",ylab = "Count")

# 3. Visitor type variable
visit <- df2$VisitorType
visit.freq <- table(visit)
barplot(visit.freq,col = "pink",xlab = "Visitor type",ylab = "Count")

# 4. Month variable
month<- df2$Month
month.freq <- table(month)
barplot(month.freq,col = "orange",xlab = "Month",ylab = "Count")

# 5. Operating systems
os<- df2$OperatingSystems
os.freq <- table(os)
barplot(os.freq,col = "blue",xlab = "OS",ylab = "Count")

# 6. Browser
brow<- df2$Browser
brow.freq <- table(brow)
barplot(brow.freq,col = "maroon",xlab = "Browser",ylab = "Count")

# 7. Administrative
adm<- df2$Administrative
adm.freq <- table(adm)
barplot(adm.freq,col = "red",xlab = "Administrative page visits",ylab = "Count")
str(df2)
# 8. Region
reg<- df2$Region
reg.freq <- table(reg)
barplot(reg.freq,col = "brown",xlab = "Region",ylab = "Count")

# 9. Informational          
info<- df2$Informational          
info.freq <- table(info)
barplot(info.freq,xlab = "Informational page visits",ylab = "Count")

# 10. TrafficType  
traff<- df2$TrafficType          
traff.freq <- table(traff)
barplot(traff.freq,xlab = "Traffic type",ylab = "Count")

# 11. Special day
special <- df2$SpecialDay
special.freq <- table(special)
barplot(special.freq,xlab = "Special day",ylab = "Count")
```

- There are more False revenue values than true revenue ones. More people didn't bring in revenue as compared to those who did from perusing the site.
- More people visited the site on weekdays compared to a lesser number on the weekends.
- Returning visitors are more compared to other types.
- May has the highest visits while February has the lowest.
- Most visitors have type 2 operating system.
- Browser 2 is used by many visitors.
- Region 1 has highest visitors while 5 has the lowest.
- Traffic type shows from what source are the visitors accessing the site from. Type 2 has most visitors.
- Most visitors did not visit the administrative and informational pages in a session.
- Most visitors visit when ther's no special day upcoming.

### 5.2.) Bivariate
```{r}
# 2 categorical variables

library(ggplot2)
ggplot(df2, aes(x = Revenue, fill = Weekend )) + 
   geom_bar(position = "dodge")


ggplot(df2, aes(x = Revenue, fill = VisitorType )) + 
   geom_bar(position = "dodge")


ggplot(df2, aes(x = Weekend, fill = VisitorType )) + 
   geom_bar(position = "dodge")

ggplot(df2, aes(x = Month, fill = Revenue )) + 
   geom_bar(position = "dodge")

ggplot(df2, aes(x = Month, fill = VisitorType )) + 
   geom_bar(position = "dodge")

# categorical and numerical

ggplot(df2, aes(Revenue, Administrative)) +
    geom_bar(stat = "identity") + 
    labs(y = "Visits on Administrative page", x = "Revenue status")

ggplot(df2, aes(Revenue, Informational)) +
    geom_bar(stat = "identity") + 
    labs(y = "Visits on Informational page", x = "Revenue status")

ggplot(df2, aes(Revenue, BounceRates)) +
    geom_bar(stat = "identity") + 
    labs(y = "Bounce rates", x = "Revenue status")

ggplot(df2, aes(Revenue, ExitRates)) +
    geom_bar(stat = "identity") + 
    labs(y = "Exit rates", x = "Revenue status")

ggplot(df2, aes(Revenue, ProductRelated_Duration)) +
    geom_bar(stat = "identity") + 
    labs(y = "Time spent on product related page", x = "Revenue status")

ggplot(df2, aes(Revenue, PageValues)) +
    geom_bar(stat = "identity") + 
    labs(y = "Page values", x = "Revenue status")

# between continuous variables

ggplot(df2, 
       aes(x = Administrative_Duration, 
           y = BounceRates)) +
  geom_point()
cor(df2$Administrative_Duration,df2$BounceRates)

ggplot(df2, 
       aes(x = Administrative_Duration, 
           y = ExitRates)) +
  geom_point()
cor(df2$Administrative_Duration,df2$ExitRates)


ggplot(df2, 
       aes(x = Administrative_Duration, 
           y = PageValues)) +
  geom_point()
cor(df2$Administrative_Duration,df2$PageValues)


ggplot(df2, 
       aes(x = Administrative_Duration, 
           y = Informational_Duration)) +
  geom_point()
cor(df2$Administrative_Duration,df2$Informational_Duration)


ggplot(df2, 
       aes(x = Informational_Duration, 
           y = BounceRates)) +
  geom_point()
cor(df2$Informational_Duration,df2$BounceRates)


ggplot(df2, 
       aes(x = Informational_Duration, 
           y = ExitRates)) +
  geom_point()
cor(df2$Informational_Duration,df2$ExitRates)


ggplot(df2, 
       aes(x = Informational_Duration, 
           y = PageValues)) +
  geom_point()
cor(df2$Informational_Duration,df2$PageValues)


ggplot(df2, 
       aes(x = Informational_Duration, 
           y = ProductRelated_Duration)) +
  geom_point()
cor(df2$Informational_Duration,df2$ProductRelated_Duration)


ggplot(df2, 
       aes(x = ProductRelated_Duration, 
           y = BounceRates)) +
  geom_point()
cor(df2$ProductRelated_Duration,df2$BounceRates)


ggplot(df2, 
       aes(x = ProductRelated_Duration, 
           y = ExitRates)) +
  geom_point()
cor(df2$ProductRelated_Duration,df2$ExitRates)


ggplot(df2, 
       aes(x = ProductRelated_Duration, 
           y = PageValues)) +
  geom_point()
cor(df2$ProductRelated_Duration,df2$PageValues)

```

- More people bring in revenue on weekdays compared to weekends.
- New visitors who do not bring in revenue are higher than the new ones who do. Same goes for returning visitors. However a higher number of returning visitors bring in revenue compared to new visitors.
- Returning visitors are more than new visitors both on weekdays and weekends.
- November has the highest revenue while February has the lowest.
- May has the highest number of returning visitors while Feb has the lowest. November has the highest number of new visitors while Feb has the lowest.
- Number of visits on administrative, Informational page translating into revenue are lower than those that don't bring in revenue.
- A higher percentage that was the last in the session translates into no revenue compared to a lower percentage that generates revenue.
- More time spent on product related page doesn't necessarily translate into revenue.
- Exit rates and bounce rates are low for visitors that generate revenue while they are higher for visitors that don't generate revenue.
- Page value is higher in visitors that produce revenue compared to those who don't.
- Exit rates and bounce rates are positively correlated, meaning an increase in one leads to increase in another.
- Exit rates is negatively correlated with duration spent on product page, hence an increase in duration leads to decrease in exit rate and vice versa.

## 6.)Implement the Solution

### 6.1) K-Means Clustering

```{r}
a <- names(df2)[sapply(df2, is.integer)]
a

df3 <- df2[,-c("Revenue", "Administrative", "Informational", "ProductRelated", "OperatingSystems"
, "Browser", "Region", "TrafficType","Month","Weekend","VisitorType")] # exctracting numerical data only
df3.class<- df2[, c("Revenue","Month","Weekend","VisitorType")]
head(df3)
head (df3.class)
str(df3)

# Normalization
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}

df3 <- normalize(df3)
head(df3)

# Applying the K-means clustering algorithm with (k)=2

kmeans <- kmeans(df3,2) 
kmeans

kmeans$centers # Getting the value of cluster center data point value(2 centers for k=2)

kmeans$cluster # showing distribution of members per cluster

plot(df3[,3:4], col = kmeans$cluster) # shows how data in features bounce rates and product related duration is split into 2 clusters
```

-2 clusters of sizes 11268, 931.
- sample 2 has the lowest distance(1.887334) between it's components while sample 1 has the highest distance(3.354206) between components.

We perform this exercise in a loop to find updated cluster centers and allocation of each observation. The iteration will stop when the maximum number of iterations is reached or the change of within-cluster sum of squares in two successive iterations is less than the threshold value. The updated cluster centers for the last iteration are called Final Cluster Centers.

- 53% is a measure of the total variance in the data set that is explained by the clustering. k-means minimizes the within group dispersion and maximizes the between-group dispersion. By assigning the samples to k clusters rather than n (number of samples) clusters achieved a reduction in sums of squares of 53%.

_ K-Means Clustering was advantageous to use as it was easy to implement.However, it is
sensitive to scale and since we normalized the dataset,it completely changed results. 


# 6.2) Hierarchical Clustering
```{r}
head(df3)

# Computing the Euclidean distance between observations using dist()

b <- dist(df3, method = "euclidean")
#print(b,digits=3)

# Hierarchical clustering. By default, it uses complete linkage
complete.hc <- hclust(b,method = "ward.D2")

# Plotting the dendrogram

plot(complete.hc,cex = 0.6, hang = -1)

# checking cluster membership
complete.member <- cutree(complete.hc,2)
table(complete.member)
```

- No need to specify the number of clusters is an advantage of this cluster type. It is also easy to implement. 
- When clustered into 2, 9489 components cluster in 1 while 2710  are in 2.

## 7.)Challenge the Solution

- Hierarchical clusters better in this case compared to kmeans when k=2

## 8.)Follow up Questions

- What is needed to be done in order for classifications to be more precise.
